% Remerciements / Acknowledgements
%
%  Grâce aux remerciements, l'auteur attire l'attention du lecteur
% sur l'aide que certaines personnes lui ont apportée, sur leurs
% conseils ou sur toute autre forme de contribution lors de la
% réalisation de son mémoire. Le cas échéant, c'est dans cette section
% que le candidat doit témoigner sa reconnaissance à son directeur de
% recherche, aux organismes dispensateurs de subventions ou aux
% entreprises qui lui ont accordé des bourses ou des fonds de
% recherche.
\ifthenelse{\equal{\Langue}{english}}{
	\chapter*{ACKNOWLEDGEMENTS}\thispagestyle{headings}
	\addcontentsline{toc}{compteur}{ACKNOWLEDGEMENTS}
}{
	\chapter*{REMERCIEMENTS}\thispagestyle{headings}
	\addcontentsline{toc}{compteur}{REMERCIEMENTS}
}
%

First, I would like to thank Professor Giovanni Beltrame for giving me a chance to join the MistLab two years ago and pursue a master's degree in robotics. Coming from an aerospace engineering background, I had no significant knowledge of this field and there was no guarantee that my involvement in the laboratory would turn out fruitful. I am grateful to have received this opportunity, the last two years have been immensely enriching.

I would also like to thank Samuel Arseneault with whom I had the chance to spend a lot of time over the last two years. His coding meticulousness and work ethic have been of tremendous help in the various research projects we have worked on.

I must also thank Polytechnique Montréal, the Natural Sciences and Engineering Research Council of Canada and Mitacs for providing financial support which allowed me to pursue my master's degree.

Finally, I would like to thank my family and friends to have made the last two years enjoyable even in the context of the pandemic. Having activities to look for beside the master's degree definitely helps keep a more balanced lifestyle.
